Social Network Analysis: Week 3: Degree and Centrality

### 2022-02-22 docs/0. Setting UP Project/Week0_About_Distill-Blog.Rmd

---
title: "Social Network Analysis: Week 3: Degree and Centrality"
description: " "
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "Umass Amherst - DACSS 679: Social Network Analysis"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

@@

## *Briefly describe Smuggler's Cove Drink Index dataset:*

@@

:   is a legendary cocktail bar in San Francisco.
    Since opening in 2009, this Tiki Bar has
    developed a dedicated following and received
    many awards. What do people in the restaurant
    industry do when they receive any modicum of
    celebrity?

@@

The network data contained in a cocktail recipe
book can serve as a proxy for some of the more
traditional elements of social network analysis.

@@

> *How do the social facts of and author's food
> culture shape their recipe selection?*
>
> The author's of[Smuggler's
> Cove](https://www.smugglerscovesf.com/#about-section)
> could not have opened their bar in a place where
> the community's alcohol food taboo is codified
> in the law, which could be many places - all of
> which are not San Francisco. Since the authors
> did live in California during the early 21st
> century they were exposed to the 21st century
> cocktail culture revival. The ability for the
> authors to write about [Exotic Cocktails, Rum,
> and the Cult of
> Tiki](https://www.amazon.com/Smugglers-Cove-Exotic-Cocktails-Cult/dp/1607747324)
> is wholly contingent upon recreational
> consumption of the alcohol being a deeply
> ingrained element of San Francisco's food
> culture - which developed a cocktail focused
> sub-culture.
>
> When "Prohibition" was a social fact in the
> early 20th century U.S.A. the legally codified
> alcohol taboo was an element of the larger
> culture's class stratification and racial cast
> system. The perpetrators of Temperance Movement
> restricted alcohol access as a means of social
> control - specifically, the socially
> undesirable. While their social betters enjoyed
> the privileged of maintaining (extra)legal
> access to alcohol. This period's cocktail
> culture being determined by access to alcohol
> could be used as a proxy for examining proximity
> to political power during the Temperance
> Movement. Besides the venn diagram of
> communities target by the Temperance Movement
> and the target populations of nascent eugenics
> movement quite possibly being just one circle -
> elements of cocktail culture can serve as a
> proxy to explore network dynamics of early 20th
> pre-globalized international markets, in
> addition to agricultural products immediately
> before The Dust Bowl and advent of
> refrigeration.

@@

![](https://tenor.com/brZQ5.gif){width="2474"
height="255"}

@@

This isn't the "ask 10 people and get 10 different
answers" sort of question, but all of the answers
are based on each person's professionally
(un)informed opinion. As a former bartender - I
believe in the authority of the published opinions
of professional bartenders and [cocktail
historians](https://www.theatlantic.com/health/archive/2010/06/how-to-be-a-drink-historian/58724/).
Their opinion is based in direct experience and
observation of the subject in question. Which is
the sort of thing empiricists do, but since they
are building cocktails instead of a dataset, their
answers are given the same consideration of
qualitative research.

@@

## I'm curious to see what sort of relationships exists between Tiki Cocktail Ingredients?

## Cocktails are build with specific ingredients, methods, and presentations to fit contemporary cocktail culture aesthetics.

### *Write up your results briefly:*

-   focus on interpretation of the results.
-   Don't list statistics or provide code,
-   **the goal is to choose a few interesting
    results and provide an interpretation.**

### Working from the example R syntax:

1.  calculate network density;

2.  degree centrality, and

3.  centralization measures for your network data.

<div>

> -   identify initial network format;
>
>     -   **Smuggler's Cove Index was converted
>         into an edgelist**
>
> -   describe and identify the nodes including:
>
>     -   **Smuggler's Cove Edgelist**
>
>         -   **Nodes: 274**
>
>         -   **Edges: 672**
>
> -   what constitutes a tie or edge including:
>
>     -   Ties
>
>         -   Number:
>
>         -   Directed
>
>         -   Binary
>
>     -   interpret the value of the tie if any.
>
>     -    edge attributes that might be used to:
>
>         -   subset data, or
>         -   stack multiple networks.

</div>

## *Provide at least two or three noteworthy results*, including:

1.  the relevant statistics and interpretation.

    -   For example, explaining which **node(s)
        are most central and which are least
        central.**

2.  Discuss (with any related evidence) whether or
    not the **node(s) behavior is in line with or
    violates expectations based on the degree
    centrality measure.**

    1.  **What do you make of network density and
        centralization measures?\
        **

        +-----------+--------------+--------+-----------+
        | Analysis  | statnet      | igraph | Plot      |
        +===========+==============+========+===========+
        | Dyad      | Mutual: 1    | 0.0179 | yes       |
        | Census    |              |        |           |
        |           | A            |        |           |
        |           | symmetrical: |        |           |
        |           | 670          |        |           |
        |           |              |        |           |
        |           | Null: 36730  |        |           |
        +-----------+--------------+--------+-----------+
        | Triad     | 0: 3216391   |        |           |
        | Census:   |              |        |           |
        |           | 1: 7879 0    |        |           |
        | All       |              |        |           |
        | Triads in | 2 : 166754   |        |           |
        | Network   |              |        |           |
        |           | 3:           |        |           |
        +-----------+--------------+--------+-----------+
        | Triad     | 3391024      | (undi  | yes       |
        | Census:   |              | rected |           |
        |           |              | graph) |           |
        | total     |              |        |           |
        | number of |              | [1]    |           |
        | all four  |              | 3      |           |
        | triad     |              | 216659 |           |
        | types     |              |        |           |
        |           |              | [2]    |           |
        |           |              | 166486 |           |
        |           |              |        |           |
        |           |              | [11]   |           |
        |           |              | 7879   |           |
        +-----------+--------------+--------+-----------+
        | Tra       | undirected   | Undi   |           |
        | nsitivity |              | rected |           |
        +-----------+--------------+--------+-----------+
        | Local     |              |        |           |
        | tran      |              |        |           |
        | sitivity: |              |        |           |
        | Node List |              |        |           |
        | & Ego     |              |        |           |
        | Network   |              |        |           |
        +-----------+--------------+--------+-----------+
        | Isolates  | no           |        |           |
        +-----------+--------------+--------+-----------+
        | Un        |              | n      |           |
        | directed: |              | umber: |           |
        | Component |              | 1      |           |
        | Structure |              |        |           |
        |           |              | size:  |           |
        |           |              | 274    |           |
        +-----------+--------------+--------+-----------+
        | Directed: |              | 0.01   | yes       |
        | GRAPH:    |              | 671642 |           |
        | Dyad      |              |        |           |
        | Census    |              |        |           |
        +-----------+--------------+--------+-----------+
        |           |              |        | yes       |
        | Directed: |              |        |           |
        | Degree    |              |        |           |
        | C         |              |        |           |
        | entrality |              |        |           |
        +-----------+--------------+--------+-----------+
        | Directed: |              |        | yes       |
        | Be        |              |        |           |
        | tweenness |              |        |           |
        | C         |              |        |           |
        | entrality |              |        |           |
        +-----------+--------------+--------+-----------+
        | Directed: |              |        | yes       |
        | Ei        |              |        |           |
        | genvector |              |        |           |
        | C         |              |        |           |
        | entrality |              |        |           |
        +-----------+--------------+--------+-----------+
        | Directed: |              |        |           |
        | Bonacich  |              |        |           |
        | Ce        |              |        |           |
        | ntrality} |              |        |           |
        +-----------+--------------+--------+-----------+
        |           |              |        |           |
        | Directed: |              |        |           |
        | Page Rank |              |        |           |
        +-----------+--------------+--------+-----------+
        |           |              |        |           |
        +-----------+--------------+--------+-----------+
        |           |              |        |           |
        +-----------+--------------+--------+-----------+

> Distill is a publication format for scientific
> and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.

>>>>

---
title: "Social Network Analysis: Week 0: About this dataset"
description: " "
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "UMass Amherst: DACSS 679: Social Network Analysis: Assignments"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

@@

## *About the Smuggler's Cove Drink Index dataset:*

@@

:   is a legendary cocktail bar in San Francisco.
    Since opening in 2009, this Tiki Bar has
    developed a dedicated following, received many
    awards, and is high-key famous among cocktail
    corps d'elite.

![](https://media.giphy.com/media/l2JJIQmEvS4hBbnjO/giphy.gif)

[Exotic Cocktails, Rum, and the Cult of Tiki](https://www.smugglerscovesf.com/#book-section)

:   What do people in the restaurant industry do
    when they receive any modicum of celebrity?

@@

Converting the data and metadata in recipe book
into network data could serve as a proxy to
reflect on more conventional network datasets.
Researchers into business and profitability could
recipe books cum network data to perform
historical market research - but all of those
people are not me.

@@

How do the social facts of and author's food culture shape their recipe selection?

:   The author's of[Smuggler's
    Cove](https://www.smugglerscovesf.com/#about-section)
    could not have opened their bar in a place
    where the community's alcohol food taboo is
    codified in the law, which could be many
    places - all of which are not San Francisco.
    Since the authors did live in California
    during the early 21st century they were
    exposed to the 21st century cocktail culture
    revival. The ability for the authors to write
    about [Exotic Cocktails, Rum, and the Cult of
    Tiki](https://www.amazon.com/Smugglers-Cove-Exotic-Cocktails-Cult/dp/1607747324)
    is wholly contingent upon recreational
    consumption of the alcohol being a deeply
    ingrained element of San Francisco's food
    culture - which developed a cocktail focused
    sub-culture.

    When "Prohibition" was a social fact in the
    early 20th century U.S.A. the legally codified
    alcohol taboo was an element of the larger
    culture's class stratification and racial cast
    system. The perpetrators of Temperance
    Movement restricted alcohol access as a means
    of social control - specifically, the socially
    undesirable. While their social betters
    enjoyed the privileged of maintaining
    (extra)legal access to alcohol. This period's
    cocktail culture being determined by access to
    alcohol could be used as a proxy for examining
    proximity to political power during the
    Temperance Movement. Besides the venn diagram
    of communities target by the Temperance
    Movement and the target populations of nascent
    eugenics movement quite possibly being just
    one circle - elements of cocktail culture can
    serve as a proxy to explore network dynamics
    of early 20th pre-globalized international
    markets, in addition to agricultural products
    immediately before The Dust Bowl and advent of
    refrigeration.

@@

Cocktails are build with specific ingredients,
methods, and presentations to fit contemporary
cocktail culture aesthetics.

![](https://media3.giphy.com/media/zbW5zoQetefZwdBR6Z/200w.webp?cid=ecf05e477of2cy4i1vdat12zy8n4ufrcuuikyv6qxkmwe1z2&rid=200w.webp&ct=g){width="400"}

@@

This isn't the "ask 10 people and get 10 different
answers" sort of question, but all of the answers
are based on each person's professionally
(un)informed opinion. As a former bartender - I
believe in the authority of the published opinions
of professional bartenders and [cocktail
historians](https://www.theatlantic.com/health/archive/2010/06/how-to-be-a-drink-historian/58724/).

Their opinion is based in direct experience and
observation of the subject in question. Which is
the sort of thing empiricists do, but since they
are building cocktails instead of a dataset, their
answers are given the same consideration of
qualitative research.

@@

I'm curious to see what sort of relationships
exists between cocktail ingredients?

![](C:/Users/white/AppData/Local/RStudio/tmp/paste-FF879F11.png){width="500"}

> `D`istill is a publication format for scientific
> and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.

### 2022-02-22 docs/3. SNA - Assignments/Week3_DRAFT_Results.Rmd

      *Week 3 Assignment: Degree and Centrality*

Using the tutorial as a guide, 
      *  you may use your own data; 
      *  the data provided in the week 3. Syntax dataset is Correlates of War trade network - you can read more about it online; or
      *  choose another dataset.

## Working from the example R syntax:
1.   calculate network density; 
2.   degree centrality, and 
3.   centralization measures for your network data.

## Write up your results briefly:
*   focus on interpretation of the results. 
*   Don’t list statistics or provide code, 
*  **the goal is to choose a few interesting results and provide an interpretation.**

## Briefly describe the dataset:
1.   identify initial network format; 
2.   describe and identify the nodes 
     including: 
        *  how many nodes are in the dataset; 
3.   what constitutes a tie or edge
     including:
        *  how many ties, 
        *   whether ties are directed/undirected,
        *   whether ties are weighted/binary, 
        *   and how to interpret the value of the tie if any.
        *   Whether or not there are edge attributes 
            that might be used to:
            *   subset data, or 
            *   stack multiple networks. 
Not every feature of the network needs to be described, but **description should orient reader to the network data and provide any necessary context for the results provided.**

## Provide at least two or three noteworthy results, including:
1.   the relevant statistics and interpretation. 
        For example, explaining which node(s) are most central and which are least central. 
2.   Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure. 
3.   What do you make of network density and centralization measures?  
4.   What about differences in the binary 100 million threshold network vs valued trade flows network. 
 5.  Check tutorial for other ideas of results to discuss.

##
1.  [**Identify an existing data set.**]{.ul}

| This can be one provided in the course directory, in an R package or library, located online, or some other source.

2.  [**Briefly describe the network dataset.**]{.ul}

| Identify and describe content of nodes and links, and identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and whether attribute data are present. 
  Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)

# Week 2: Scripts

```{r Import Data Set, echo=TRUE}

  # install.packages("network")

    data("flo", package = "network")

# 2. Briefly describe the network dataset.

-   *Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)*

1.  *Identify and describe content of nodes and links,*

-   each of the 16 Florentine Family Names represent a Node in the network.

```{r Colnames}
 colnames(flo)
```

2.  *Identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and*

    -   Flo Format: Size: 16 x 16 square dimmentions indicate a adjacency matrix format

    Network attributes: vertices = 16 directed = TRUE hyper = FALSE loops = FALSE multiple = FALSE bipartite = FALSE total edges= 40 missing edges= 0 non-missing edges= 40

Vertex attribute names: vertex.names

No edge attributes

```{r Dimmentions, echo=TRUE}

  dim(flo)

```{r Network Attributes, echo=TRUE}

  network(flo)

3.  *Identify whether attribute data are present.*

```{r Summary,echo=TRUE}

  summary(flo,
          print.adj = TRUE
          )
```

## Create & Describe: Directed Adjacency Matrix

*Florentine Families Directed Adjacency Matrix Network Objects:*
    **iGraph** flo.ig 
    **Statnet** flo.stat

iGraph and Statnet both use {r graph4} for Directed Adjacency Matrix Objects

```{r graph4, statnet: Directed: Adjacency Network, echo=TRUE}

    flo.stat<-
        network(
        flo,
        directed=F,
        matrix.type="adjacency"
        )

#     provides a description of several critical network features
    print(flo.stat)

```{r graph4, igraph: Directed: Adjacency Network, echo=TRUE}

    flo.ig<-
      graph.adjacency(
      flo,
      mode="undirected"
      ) 
    print(flo.ig)
```

```{r igraph: Network Description, echo=TRUE}

#     Count Vertices
    vcount(flo.ig)

#     Count Edges
    ecount(flo.ig)

# Is this a Bipartite or single mode network?
    is_bipartite(flo.ig)

#  Are edges directed or undirected?
    is_directed(flo.ig)

#Are edges weighted or unweighted?
    is_weighted(flo.ig)

```
 **Vertex and Edge Attributes**

```{r statnet: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attributes
  head(flo.stat %v% "vertex.names")

#      list the names of vertex attributes
  network::list.vertex.attributes(flo.stat)

#     access edge attribute
  head(flo.stat%e% "weight")

#  Generate a list the names of edge attributes
  network::list.edge.attributes(flo.stat)

#   summarize numeric network attribute
  summary(flo.stat  %v% "name")

```{r igraph: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attribute 
  V(flo.ig)$name
#     Generate a list the names of vertex attributes
  igraph::vertex_attr_names(flo.ig)

#     access edge attribute
  E(flo.ig)$weight

#   Generate a list the names of edge attributes  
  igraph::edge_attr_names(flo.ig)

#   summarize numeric network attribute
  summary(E(flo.ig)$weight)

# 3. Explore the dataset using commands from week 2 tutorial.

5.  *geodesic and path distances,*

## Path Length and Geodesic

```{r iGraph:Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(flo.ig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(flo.ig,"Bischeri", "Castellani",weights=NA)

We can also find all of the shortest paths between two famalies in this network.
These list the starting node and the ending node with all nodes inbetween for each path. 
The option weights=NA means that any available edge weights are ignored.

```{r iGraph:Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(flo.ig,"Strozzi","Tornabuoni", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

The concept of shortest path to describe the overall network structure can be more useful

```{r iGraph:Network: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(flo.ig,directed=F)

6.  *triads or transitivity,*

## Dyad Census

Classifies all dyads in the network as: - Reciprocal (mutual), or mut - Asymmetric (non-mutual), or asym, and - Absent, or null

```{r statnet: Dyad Census , echo=TRUE}

    sna::dyad.census(flo.stat)

```{r iGraph: Dyad Census, echo=TRUE}

  igraph::dyad.census(flo.ig)

## Triad Census

Classifies all triads in the network. The triad census provides a fundamental descriptive insight into the types of triads found in a particular dataset.

```{r statnet: Triad Census, echo=TRUE}

#     Classify all Triad in the network: Directed Network
  sna::triad.census(flo.stat, mode= )

#     Classify all Triad in the network: Undirected Network
   sna::triad.census(flo.stat, mode="graph")

#     total number of all four triad types returned by triad census
   sum(sna::triad.census(flo.stat, mode="graph"))

```{r iGraph: Triad Census, echo=TRUE}
#Classify all Triad in the network
     igraph::triad_census(flo.ig)

## Transitivity or Global Clustering

```{r iGraph: weighted/ network transitivity, echo=TRUE}

#     network transitivity:
    transitivity(flo.ig)

#     weighted network transitivity:
#   igraph::transivity(flo.ig)

```{r statnet: weighted/ directed network transitivity, echo=TRUE}

#     Directed Networks transitivity:
  gtrans(flo.stat)

#   methods for weighted networks, rank and correlation, along with the relevant references.
#  sna::gtrans()

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.

```{r iGraph:Local transitivity, echo=TRUE}

#     Retrive a list of the vertices we are interested
    V(flo.ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(flo.ig,
             type="local", 
             vids=V(flo.ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(flo.ig, type="global")

#     get average local clustering coefficient
  transitivity(flo.ig, type="average")

7.  *connectedness and.or component structure, etc.*

> Distill is a publication format for scientific and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
##

>>>>

## Working from the example R syntax:

1.  calculate network density;
2.  degree centrality, and
3.  centralization measures for your network data.

## *Write up your results briefly:*

-   focus on interpretation of the results.
-   Don't list statistics or provide code,
-   **the goal is to choose a few interesting
    results and provide an interpretation.**

## *Briefly describe the dataset:*

# 1. identify initial network format

<div>

> -   identify initial network format;
>
>     -   **Smuggler's Cove Index was converted
>         into an edgelist**
>
> -   what constitutes a tie or edge including:
>
>     -   Ties
>
>         -   Number:
>
>         -   Directed:
>
>         -   Binary:
>
>     -   interpret the value of the tie if any.
>
>     -   edge attributes that might be used to:
>
>     <!-- -->
>
>         -   subset data, or
>         -   stack multiple networks.

</div>

# 2. describe and identify the nodes including:

<div>

**Smuggler's Cove Edge-list contains:**

> -   Nodes: 274
>
>     -   Nodes Values:
>
>         -   [Drink Name]{.smallcaps}
>
>         -   [Drink Ingredient]{.smallcaps}
>
> -   Edges: 672
>
>     -   Drink Ingredient Type

</div>

# 3. what constitutes a tie or edge including:

-   how many ties,

-   whether ties are directed/undirected,

-   whether ties are weighted/binary,

-   and how to interpret the value of the tie if
    any.

-   Whether or not there are edge attributes that
    might be used to:

    -   subset data, or
    -   stack multiple networks.

# *Provide at least two or three noteworthy results*, including:

## 1. the relevant statistics and interpretation. For example, explaining which node(s) are most central and which are least central.

## 2. Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure.

## 3.What do you make of network density and centralization measures?

## 4. What about differences in the binary 100 million threshold network vs valued trade flows network.

## 5. Check tutorial for other ideas of results to discuss.

> Distill is a publication format for scientific
> and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.
